---
title: "An Example of DiagnosisInformation"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mengbing/Box Sync/OptumInsight_DataManagement/data_freeze/20181020/data_icd9TreeStructured')

library(dplyr)
library(tidyr)
library(data.table)
library(reshape2)
library(data.tree)
library(Matrix)
library(knitr)
```

This file shows an example of diagnosisInformation.RData, output from data_icd9TreeStructured_list.R. 
  
We consider both inpatient and outpatient diagnoses from medical claims data, obtained from rows in `diagData_20181020_freeze.csv` that have `source %in% ("medical.inpatient", "medical.outpatient")`. The diagnoses range from individual index VTE dates to 2015-09-30. All diagnoses, regardless of being before or after index VTE date, are included. The `diagnosisInformation_list.RData` contains a list named diagnosisInformation of 14,883 patients. 

Patient counts:

- 14,945 = total number of patients in our study cohort. These patients satisfy all inclusion/exclusion criteria. 
- 14,883 = number of patients in our study cohort who had at least one diagnosis in medical claims before 2015-09-30. The 14945 - 14883 = 62 excluded patients did not have any diagnoses in medical claims before 2015-09-30, although they might have outpatient pharmacy claims after 2015-10-01.
- 14,932 = number of patients in our study cohort who had at least one diagnosis in either medical or confinement claims before 2015-09-30. The 14932 - 14883 = 49 patients had only confinement claims but no medical claims before 2015-09-30.

Some diagnosis codes before 2015-09-30 are also ICD-10. This is strange.

The list contains:

- _patients_ = a character vector of patient ids
- _roots_ = a character vector of roots of the ICD-9 trees, i.e. 3-digit ICD-9 codes. roots[i] is a 3-digit ICD-9 code of the patient with ID patid[i]
- _trees_ = a list of sparse adjacency matrices that represent the tree-structured ICD-9 codes. The name of each element is the root of the tree.
- _nodes_ = a list of character vectors containing the nodes of each tree in trees. The name of each element is the root of the tree.
- _counts_ = a list of numeric vectors containing the number of times each node appears
in the patient's medical record, regardless of date of service. The name of each element is of the form "patid:rootName".

We show an example of patient 802666500100523.

The 3-digit ICD-9 codes that patient 802666500100523 had at least once in medical claims data.
```{r}
load("diagnosisInformation_list.RData")
testIndex <- which(diagnosisInformation$patients=="802666500100523")
diagnosisInformation$roots[testIndex]
```

Now we look at one of the trees with root "388". The tree is represented using an adjacency matrix. Denote this matrix as $M$. $M_{jk} = 1$ if the $k$-th node is a child of the $j$-th node. Hence, each rowSum equals the number of children the corresponding node has. Each column has at most one 1.
```{r}
diagnosisInformation$trees$`388`
```

Here is the vector of all nodes in the tree with root "388". This vector contains the row names of the sparse matrix above.
```{r}
diagnosisInformation$nodes$`388`
```

The _counts_ vector counts how many times patient "802666500100523" was diagnosed with each node in the tree with root "388". The counting follows the rule that if a 4/5-digit ICD-9 code is present (node = 1), then the corresponding 3/4-digit ICD-9 code is present.
```{r}
diagnosisInformation$counts$`802666500100523:388`
```



Here are some diagnosis codes that are not available from the ICD-9 code dictionary.
```{r}
diagnosisInformation$nodes$`198`
diagnosisInformation$nodes$`174`
diagnosisInformation$nodes$`401`
```


I would like to look at which patient has "4019Y".
```{r results='asis'}
n <- length(diagnosisInformation$nodes$`401`)
indx <- which(diagnosisInformation$roots == "401")
patientIndx <- diagnosisInformation$patients[indx]
countIndx <- diagnosisInformation$counts[indx]

patientsToPrint <- c()
for (j in 1:length(patientIndx)){
  diagCounts <- diagnosisInformation$counts[indx[j]]
  if (diagCounts[[1]][n] > 0) {
    patientsToPrint <- c(patientsToPrint,
                         diagnosisInformation$patients[indx[j]])
  }
}
patientsToPrint
```










