---
title: "test_codes_MCMCglmm"
output: html_notebook
---

* Reference: https://cran.r-project.org/web/packages/MCMCglmm/vignettes/CourseNotes.pdf 

# Section 2.6 Categorical Data

```{r}
# simulate 2 binomial variables each with 5 trials
success <- rbinom(2, size = 5, prob = c(0.4,0.6))
failure <- 5 - success
binom <- rbind(success, failure)
colnames(binom) <- c("u1", "u2")
binom

# expand into success or failure
binary <- matrix(rep(c(1,0,1,0), binom), 1, 10)
colnames(binary) <- rep(c("u1", "u2"), each=5)
binary
```

Pupated variable is zero if they failed to pupate or one if they successfully pupated. The 874 individuals are spread across 49 full-sib families, with family sizes ranging from 6 to 38.
```{r}
library(MCMCglmm)
PlodiaRb <- data("PlodiaRB")
head(PlodiaRb)
```


## Fix the residual variance at 1:
```{r}
prior.m2b.1 <- list(R = list(V = 1, fix = 1), G = list(G1 = list(V = 1, nu = 0.002)))

m2b.1 <- MCMCglmm(Pupated ~ 1, random = ~FSfamily, family = "categorical", data = PlodiaRB, prior = prior.m2b.1, verbose = FALSE)

summary(m2b.1)
```

## Fix the residual variance at 2:
```{r}
prior.m2b.2 <- list(R = list(V = 2, fix = 1), G = list(G1 = list(V = 1, nu = 0.002)))

m2b.2 <- MCMCglmm(Pupated ~ 1, random = ~FSfamily, family = "categorical", data = PlodiaRB, prior = prior.m2b.2, verbose = FALSE)

summary(m2b.2)
```

Plot the posterior distribution for the intercept of the two models:
```{r}
plot(mcmc.list(m2b.1$Sol, m2b.2$Sol))
```

Plot the posterior distribution for the variance components of the two models:
```{r}
plot(mcmc.list(m2b.1$VCV, m2b.2$VCV))
```




