---
title: "Hospitalizations"
output:
  html_document: default
  html_notebook: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mengbing/Box Sync/OptumInsight_DataManagement/analysis')

library(dplyr)
library(tidyr)  
library(data.table)
library(knitr)
library(kableExtra)
library(xlsx)
library(readxl)
library(ggplot2)
library(reshape2)
library(pscl)
```


```{r cache=TRUE}
## Hospitalization of eligible patients
conf <- data.table(read_excel("../data/prog2_conf.xlsx"))
colnames(conf) <- tolower(colnames(conf))
conf$index_dt <- NULL
hosp_elig_cnts <- conf[, .N, by=patid]

## All hospitalizations in the confinement data, including those ineligible
conf_all <- fread("../data/conf_all.txt")
colnames(conf_all) <- tolower(colnames(conf_all))
conf_all$los <- as.numeric(conf_all$los)
conf_all$patid <- as.character(conf_all$patid)
hosp_all_cnts <- conf_all[, .N, by=patid]
hosp_cnts <- rbind(hosp_elig_cnts, hosp_all_cnts)
hosp_cnts$group <- c(rep("Eligible patients", nrow(hosp_elig_cnts)),
                     rep("All patients", nrow(hosp_all_cnts)))
```

```{r cache=TRUE}
ggplot(data=hosp_cnts[group=="Eligible patients",], aes(x=N, fill="salmon")) +
  geom_histogram(bins=30, alpha=0.2) +
  # geom_histogram(aes(y=..density..),alpha=0.2) + 
  labs(title="Histogram of number of hospitalizations of eligible patients",
       x="Number of hospitalizations") +  scale_x_continuous(limits = c(0, 30)) +
  theme(legend.position="none")
print("Eligible patients:"); summary(hosp_cnts[group=="Eligible patients", N])

ggplot(data=hosp_cnts[group=="All patients",], aes(x=N, fill="salmon")) +
  geom_histogram(bins=30, alpha=0.2) +
  # geom_histogram(aes(y=..density..),alpha=0.2) + 
  labs(title="Histogram of number of hospitalizations of all patients",
       x="Number of hospitalizations") + 
  scale_x_continuous(limits = c(0, 30)) +
  theme(legend.position="none")
print("All patients:"); summary(hosp_cnts[group=="All patients",N])
```


## Length of stay
```{r cache=TRUE}
# combine two data sets
conf2 <- rbind(conf[,.(patid, charge, copay, los)], conf_all[,.(patid, charge, copay, los)])
conf2$group <- c(rep("Eligible patients", nrow(conf)),
                 rep("All patients", nrow(conf_all)))

# negative length of stay
print("Patients with negative length of stays:"); print(conf_all[los<0,])

ggplot(data=conf2[los>=0 & group=="Eligible patients",], aes(x=los, fill="salmon")) +
  geom_histogram(binwidth = 1, alpha=0.2) +
  # geom_histogram(aes(y=..density..),alpha=0.2) + 
  labs(title="Histogram of lengths of stay of eligible patients",
       x="Length of stay") +
  scale_x_continuous(limits = c(0, 50)) +
  theme(legend.position="none")
print("Eligible patients:"); summary(conf2[group=="Eligible patients", los])

ggplot(data=conf2[los>=0 & group=="All patients",], aes(x=los, fill="salmon")) +
  geom_histogram(binwidth = 1, alpha=0.2) +
  # geom_histogram(aes(y=..density..),alpha=0.2) + 
  labs(title="Histogram of lengths of stay of all patients",
       x="Length of stay") +
  scale_x_continuous(limits = c(0, 50)) +
  theme(legend.position="none")
print("All patients:"); summary(conf2[group=="All patients" & los>=0, los])
```



```{r cache=TRUE}
dat <- data.table(readRDS("../data/prog100_analysis_data.rds"))

# calculate the number of comorbidities
comorb_names <- colnames(dat)[grep("comorbidity", colnames(dat))]
dat[, (comorb_names) := lapply(.SD, function(x) ifelse(is.na(x), 0, x)), .SDcols = comorb_names]
dat[, ("n_comorb") := rowSums(.SD), .SDcols=comorb_names]
patinfo <- unique(dat[patid %in% conf$patid, .(patid, n_comorb, cancer_type_combined)])

conf3 <- merge(x=conf2, y=patinfo, by="patid", all.x=TRUE)
conf3 <- merge(x=conf3, y=unique(hosp_cnts[,.(patid, N)]), by="patid", all.x = TRUE)
ggplot(data=conf3[los>=0 & group=="Eligible patients" & copay>0 & copay<300,], aes(x=N, y=log10(los), color=copay)) +
  geom_point() +
  scale_colour_gradient2() +
  labs(x="Number of hospitalizations", y="log10(Length of stay)",
       title="log10(Length of stay) vs # Hospitalizations \n among eligible patients by copay",
       color="Copay")
```


```{r cache=TRUE}
ggplot(data=conf3[group=="Eligible patients",], aes(x=N, y=log(los), color=n_comorb)) +
  geom_point() +
  labs(x="Number of hospitalizations", y="Length of stay",
       title="log10(Length of stay) vs log10(# Hospitalizations) \n among eligible patients among eligible patients by # of comorbidities",
       color="# comorbidities")
```




# VTE-related admissions vs non-VTE related admissions

- We only consider hospitalizations after index VTE date.

- There are 8 types of counts, depending on whether patients with "Not captured" at AC3Mo are censored:
(not censored, censored) $\times$ (n\_hosp\_vte, n\_hosp\_nonvte, n\_hosp\_vte\_associated, n\_hosp\_nonvte\_associated), where n\_hosp\_vte and n\_hosp\_nonvte are counts of hospitalizations with VTE as the primary diagnosis, and n\_hosp\_vte\_associated and n\_hosp\_nonvte\_associated are counts of hospitalizations with VTE in any of diag1 - diag5.

- Length of record of a patient is from index VTE date to the eligibility end date from member data.


## Use diag1 from confinement data as the primary reason for hospital admission.
```{r}
conf <- data.table(read_excel("../data/prog2_conf.xlsx"))
colnames(conf) <- tolower(colnames(conf))
# keep only hospitalizations after index VTE date
conf$admit_dt <- as.Date(conf$admit_dt)
conf$disch_dt <- as.Date(conf$disch_dt)
conf$index_dt <- as.Date(conf$index_dt)
conf <- conf[index_dt <= admit_dt,]

vte_codebook <- data.table(read_excel("../others/All_codes_ICD9_NDC_HCPCS.xlsx", sheet="ICD9_VTE"))
vte_codebook$ICD9_VTE <- as.character(vte_codebook$ICD9_VTE)

## Not considering censoring
# identify admissions with VTE as the primary diagnosis
icd9_vte_codes <- vte_codebook$ICD9_VTE
conf$is.vte_admission <- conf$diag1 %in% icd9_vte_codes
conf[, n_hosp_vte_prim_all := sum(is.vte_admission), by=patid]

# identify VTE-associated admission from diag1-5
is.vte_associated_bycol <- data.frame(apply(conf[, c("diag1", "diag2", "diag3", "diag4", "diag5")], 2, function(x) x %in% icd9_vte_codes))
conf$is.vte_associated <- apply(is.vte_associated_bycol, 1, any)
conf[, n_hosp_vte_associated_all := sum(is.vte_associated), by=patid]

# identify all post-indexVTE hospitalizations
conf[,  n_hosp_all := .N, by=patid]

conf[, `:=`(n_hosp_nonvte_prim_all = n_hosp_all - n_hosp_vte_prim_all,
            n_hosp_nonvte_associated_all = n_hosp_all - n_hosp_vte_associated_all)]
conf[, los_total := sum(los), by=patid]

# add indexAC and AC3Mo
dat <- data.table(readRDS("../data/prog100_analysis_data.rds"))
# combine DOACs categories
dat$index_ac2 <- with(dat, ifelse(index_ac %in% c("Apixaban", "Dabigatran", "Edoxaban", "Rivaroxaban"), "DOAC", index_ac))
dat$outcome2 <- with(dat, ifelse(outcome %in% c("Apixaban", "Dabigatran", "Edoxaban", "Rivaroxaban"), "DOAC", outcome))
comorb_names <- colnames(dat)[grep("comorbidity", colnames(dat))]
dat[, (comorb_names) := lapply(.SD, function(x) ifelse(is.na(x), 0, x)), .SDcols = comorb_names]
dat[, ("n_comorb") := rowSums(.SD), .SDcols=comorb_names]
dat2 <- unique(dat[, .(patid, index_ac_dt, index_ac, index_ac2, outcome_dt, outcome, outcome2, n_comorb)])


conf2 <- merge(x=conf, y=dat2, by="patid")


## Consider censoring: remove information after stopping date for patients with NC at 3 months

conf2[, not_censored := !(outcome2=="Not captured" & outcome_dt<=admit_dt)]
conf2[, `:=`(n_hosp_censored = sum(not_censored),
             n_hosp_vte_prim_censored = sum(not_censored*is.vte_admission),
             n_hosp_vte_associated_censored = sum(not_censored*is.vte_associated)), by=patid]
conf2[, `:=`(n_hosp_nonvte_prim_censored = n_hosp_censored - n_hosp_vte_prim_censored,
            n_hosp_nonvte_associated_censored = n_hosp_censored - n_hosp_vte_associated_censored)]
conf2[, los_total_censored := sum(los * not_censored), by=patid]

n_hosp_names <- colnames(conf2)[grep("n_hosp", colnames(conf2))]
conf3 <- unique(conf2[, (c("patid", n_hosp_names, "los_total", "los_total_censored")), with=FALSE])

## add 0 counts to patients with no hospitalization records
pat_no_conf <- dat2[!dat2$patid %in% conf3$patid, .(patid)]
pat_no_conf[, (c(n_hosp_names, "los_total", "los_total_censored")) := 0]
conf_counts <- rbind(conf3, pat_no_conf)
conf_counts <- merge(x=conf_counts, y=dat2, by="patid")
conf_counts <- merge(x=conf_counts, y=unique(dat[,.(patid, cancer_type_combined, index_vte_type)]), by="patid")

## determine length of record
member <- fread("../data/member.txt", 
                select = c("Patid", "index_dt", "Eligeff", "Eligend"),
                colClasses=list(character=1, Date=2:4, numeric=5))
colnames(member) <- tolower(colnames(member))
member$index_dt <- as.Date(member$index_dt, format = "%m/%d/%Y")
member[, 3:4] <- lapply(member[, 3:4], as.Date)

# keep enrollment periods that cover the index VTE date or start after the index VTE date
member2 <- member[index_dt <= eligend, ]
member2 <- merge(x=member2, y=dat2, by="patid")

member2$max_indexVTE_eligeff <- as.Date(apply(member2[, .(index_dt, eligeff)], 1, max))

# patients with NC at AC3Mo are censored at the stopping date
member2$min_ncdate_eligend <- as.Date(apply(member2[, .(outcome2, eligend, outcome_dt)], 1, function(x) ifelse(x[[1]]=="Not captured", min(x[[2]], x[[3]]), x[[2]])))

member2[, enroll_length_thisperiod := as.numeric(min_ncdate_eligend - max_indexVTE_eligeff)]
# length of record starts from index VTE date, and ranges over enrollment periods after index VTE date
member2 <- member2[enroll_length_thisperiod >= 0, ]
member2[, enroll_length := sum(enroll_length_thisperiod)+1, by=patid]
member2 <- member2[patid %in% conf_counts$patid,]
member3 <- unique(member2, by=c("patid", "enroll_length"))


conf_counts <- merge(x=conf_counts, y=member3[, .(patid, enroll_length)], by="patid")
conf_counts$outcome2 <- relevel(as.factor(conf_counts$outcome2), ref="LMWH")

conf_counts <- data.frame(conf_counts)
```

```{r}
ggplot(data=conf_counts, aes(x = log(n_hosp_censored), y = log(los_total_censored))) + 
  geom_point()

ggplot(data=conf_counts, aes(x = log(n_hosp_all), y = log(los_total))) + 
  geom_point() +
  geom_jitter(width = 0.2, height = 0.2, size=0.8, color="salmon", alpha=0.2)
```

The number of hospitalizations and total length of stay are positively linearly associated on log scale.



## number of patients having a gap in enrollment after index VTE date

```{r}
member4 <- member[index_dt <= eligend, ]
member4[patid %in% dat2$patid,][, N := .N, by=patid]
length(member4[N>1, patid])
```


```{r}
# check for over-dispersion
conf_counts %>% group_by(outcome2) %>%
  summarise(mean1=mean(n_hosp_vte_prim_censored), sd1=sd(n_hosp_vte_prim_censored),
            mean2=mean(n_hosp_vte_associated_censored), sd2=sd(n_hosp_vte_associated_censored)) # note that the variances are greater than the means

# check for 0 lengths
hist(member3$enroll_length, main="Histogram of enrollment length")
hist(member3$enroll_length[member3$enroll_length<=150], breaks=75, main="Histogram of enrollment length <= 150 days")
print(head(member3[enroll_length==0,]))
```


```{r}
hist(conf_counts$n_hosp_vte_prim_censored, main="Histogram of VTE hospitalizations", breaks=max(conf_counts$n_hosp_vte_prim_censored))
hist(conf_counts$n_hosp_nonvte_prim_censored, main="Histogram of VTE hospitalizations", breaks=max(conf_counts$n_hosp_nonvte_prim_censored))
```



```{r cache=TRUE, fig.width=8,fig.height=5, echo=FALSE, eval=FALSE}
### Boxplot with jittered points
ggplot(data=conf_counts, aes(x=index_ac2, y=n_hosp_nonvte_associated_censored)) +
  geom_boxplot(fill="salmon", alpha=0.7, color="salmon") +
  # geom_point(color="burlywood1") +
  geom_jitter(width = 0.3, size=0.5, color="skyblue", alpha=0.2) +
  labs(title="Boxplot of VTE-related hospitalizations by AC3Mo",
       x="AC3Mo", y="Number of hospitalizations")
```

```{r cache=TRUE, fig.width=6,fig.height=4, echo=FALSE, eval=FALSE}
ggplot(data=conf_counts, aes(x=index_ac2, y=n_hosp_vte)) +
  geom_boxplot(fill="salmon", alpha=0.7, color="salmon") +
  # geom_point(color="burlywood1") +
  geom_jitter(width = 0.3, size=0.5, color="skyblue", alpha=0.2) +
  labs(title="Boxplot of VTE-related hospitalizations by IndexAC",
       x="IndexAC", y="Number of hospitalizations")
```


### Plots of hospitalizations by outcome
```{r}
#' data should be a data.frame
#' AC_variable = "index_ac2" or "outcome2"
#' primary = TRUE if looking at admissions with VTE as the primary diagnosis, and FALSE if looking at admissions having a VTE diagnosis
#' xmax_vte_indexAC=5, xmax_nonvte_indexAC=30, xmax_vte_AC3mo=5, xmax_nonvte_AC3mo=30 are the x axis limit of histogram of VTE/non-VTE by outcome type
#' title_var is the variable that goes into the title of the plot
plot_hospitalization <- function(data, AC_variable, primary, xmax_vte_indexAC=5, xmax_nonvte_indexAC=30, xmax_vte_AC3mo=5, xmax_nonvte_AC3mo=30, title_var){ 
  AC_types <- unique(data[[AC_variable]])
  groups <- c("VTE", "Non-VTE")
  if(primary) groups_var_name <- c("n_hosp_vte_prim_all", "n_hosp_nonvte_prim_all") else
    groups_var_name <- c("n_hosp_vte_associated_censored", "n_hosp_nonvte_associated_censored")
  
  # enlarge the histogram of hospitalizations by restricting counts 
  if(AC_variable=="index_ac") his_xmax <- c(xmax_vte_indexAC, xmax_nonvte_indexAC) else
    his_xmax <- c(xmax_vte_AC3mo, xmax_nonvte_AC3mo)

  # layout where the boxplot is at top
  n <- length(AC_types)
  plot_dimension <- matrix(cbind(rep(1:2, n) + 4*rep(0:(n-1), each=2), rep(3:4, n) + 4*rep(0:(n-1), each=2)), ncol=2)
  # plot_dimension <- matrix(c(1:4), nrow=2, byrow=FALSE)
  
  # plot layout
  layout(mat = plot_dimension, height = rep(c(2,3), n))
  par(oma=c(2,2,2,2))
  
  for(AC in AC_types){
    sub_dat_byAC <- data[data[,AC_variable]==AC,]
    
    # j=1 for VTE, j=2 for non-VTE
    for(j in 1:2){
      # enlarge the histogram of non-VTE related hospitalizations by restricting counts n_hosp_non_vte to <= 30
      sub_dat <- sub_dat_byAC[sub_dat_byAC[, groups_var_name[j]] <= his_xmax[j], ]
            
      # create legends: summary of counts
      s <- summary(sub_dat[[groups_var_name[j]]])
      leg <- paste0(paste0(names(s), sep=" "), round(s,1), collapse = "\n")
    
      par(mar=c(0, 4.5, 6, 1.5))
      boxplot(sub_dat[[groups_var_name[j]]], horizontal=TRUE, ylim=c(0,max(sub_dat[[groups_var_name[j]]])), frame=F, xaxt="n", col = "salmon", main=paste0(title_var, " = ", AC, ", ", groups[j]), cex.main=1.5)
    
      par(mar=c(4, 4.5, 0, 1.5))
      hist(sub_dat[[groups_var_name[j]]], xlim=c(0,his_xmax[j]), col = "salmon", breaks=seq(0,his_xmax[j],by=1), border=F, xlab="Number of hospitalizations", ylab = "Counts", main="", cex.lab=1.5, cex.axis=1.5)
      
      legend("topright", legend=leg, col="black", box.lty=0, xpd=TRUE, bty = "n", cex=1)
    }
  }
  
  # add main title
  mtext(paste0("Number of VTE-related and non-VTE-related hospitalizations by ", title_var), outer = TRUE, cex = 1.2)
}
```


### Hospitalizations by IndexAC
```{r cache=TRUE, fig.width=8,fig.height=16}
plot_hospitalization(conf_counts, "index_ac2", primary=TRUE, title_var = "IndexAC")
```

### Hospitalizations by AC3Mo
```{r cache=TRUE, fig.width=8,fig.height=18}
plot_hospitalization(conf_counts, "outcome2", primary=TRUE, title_var = "AC3Mo")
```


## Use diag1-5 from confinement data to identify hospitalizations having an VTE.
### Hospitalizations by IndexAC
```{r cache=TRUE, fig.width=8,fig.height=16}
plot_hospitalization(conf_counts, "index_ac2", primary=FALSE, title_var = "IndexAC")
```

### Hospitalizations by AC3Mo
```{r cache=TRUE, fig.width=8,fig.height=18}
plot_hospitalization(conf_counts, "outcome2", primary=FALSE, title_var = "AC3Mo")
```





# Anticoagulant choices over time

Only days of supply is used to determine which anticoagulant a patient was on. INR was not used as an indication of warfarin use.

```{r}
dat <- data.table(readRDS("../data/prog100_analysis_data.rds"))
comorb_names <- colnames(dat)[grep("comorbidity|antiplatelet", colnames(dat))]
dat3 <- dat[, .(patid, index_dt, fill_dt, category, gen_name, index_ac, index_ac_dt, outcome, outcome_dt, days_sup)]
# combine DOACs categories
dat3$index_ac2 <- with(dat3, ifelse(index_ac %in% c("Apixaban", "Dabigatran", "Edoxaban", "Rivaroxaban"), "DOAC", index_ac))
dat3$outcome2 <- with(dat3, ifelse(outcome %in% c("Apixaban", "Dabigatran", "Edoxaban", "Rivaroxaban"), "DOAC", outcome))

# ACs of the same generic name prescribed in one day are considered , and the AC with the longest days of supply is kept
dat3 <- dat3[, .SD[days_sup==max(days_sup)], by=.(patid, fill_dt, gen_name)]
dat3$fill_days_after_indexVTE <- with(dat3, as.numeric(fill_dt - index_dt))
dat3$fill_plus_supply_after_indexVTE <- with(dat3, as.numeric(fill_days_after_indexVTE + days_sup))
max_days_after_indexVTE <- max(dat3$fill_plus_supply_after_indexVTE) #2775


#' var_name = gen_name (separate DOACs), or category (combined DOACs)
get_plot_data <- function(data, var_name){
  ac_names <- sort(unique(data[[var_name]]))
  p <- length(ac_names)
  plot_data <- data.frame(matrix(NA, nrow=max_days_after_indexVTE+1, ncol=p+1))
  plot_data[,1] <- 0:max_days_after_indexVTE
  # initialize counts
  plot_data[,-1] <- 0
  
  for(j in 1:nrow(data)){
    which_ac <- which(data[[var_name]][j] == ac_names)
    ind_start <- as.numeric(data$fill_dt[j] - data$index_dt[j])
    ind_end <- as.numeric(ind_start + data$days_sup[j]) # prescription date also counts
    # add 1 to the corresponding column of every time an AC is prescribed
    plot_data[ind_start:ind_end, which_ac+1] <- plot_data[ind_start:ind_end, which_ac+1] + 1
  }
  
  ac_names2 <- gsub("/| ", "_", ac_names)
  colnames(plot_data)[1] <- "days_after_indexVTE"
  colnames(plot_data)[1:p+1] <- paste0("n_pat_", ac_names2)
  
  # convert data into long format
  plot_data2 <- melt(plot_data, id.vars = "days_after_indexVTE")
  colnames(plot_data2)[2:3] <- c("ac_type", "n_pat")
  plot_data2$ac_type <- gsub("n_pat_", "", plot_data2$ac_type)
  
  return(plot_data2)
}

plot_ac_trend <- function(data, plot_title, plot_subtitle=NULL, max_days=NULL){
  if(!is.null(max_days)) data <- data[data$days_after_indexVTE <= max_days,]
  ggplot(data, aes(x=days_after_indexVTE)) + 
    geom_line(aes(y=n_pat, color=ac_type)) + 
    labs(x="Days after index VTE date", y="# of patients",
         title=plot_title, subtitle=plot_subtitle, color="AC types")
}
```


### Separate DOACs, over all days after index VTE date
```{r cache=TRUE}
plot_data_separate_doacs <- get_plot_data(data=dat3, var_name = "gen_name")
plot_ac_trend(data=plot_data_separate_doacs, plot_title = "# of patients on each anticoagulant over days after index VTE date", plot_subtitle="Separate DOACs")
```

### Separate DOACs, up to 90 days after index VTE date
```{r cache=TRUE}
plot_ac_trend(data=plot_data_separate_doacs, plot_title = "# of patients on each anticoagulant over up to 90 days after index VTE date", plot_subtitle="Separate DOACs", max_days = 90)
```

### Combined DOACs, over all days after index VTE date
Combined DOACs = Apixaban + Dabigatran + Edoxaban + Rivaroxaban
```{r cache=TRUE}
plot_data_combined_doacs <- get_plot_data(data=dat3, var_name = "category")
plot_ac_trend(data=plot_data_combined_doacs, plot_title = "# of patients on each anticoagulant over days after index VTE date", plot_subtitle="Combined DOACs")
```

### Combined DOACs, over up to 90 days after index VTE date
```{r cache=TRUE}
plot_ac_trend(data=plot_data_combined_doacs, plot_title = "# of patients on each anticoagulant over up to 90 days after index VTE date", plot_subtitle="Combined DOACs", max_days = 90)
```

A drastic change point at about 30 days after index VTE date occurs to all types of ACs.




# Zero-inflated Poisson regression on the number of hospitalizations by AC3Mo.

1. For patients with AC3Mo = Not captured, information after the stopping date is not used. So hospitalizations after the stopping date do not count.

2. Length of record of a patient is from index VTE date to the eligibility end date from member data.

```{r}
## Plot predicted probabilities
#' model = fitted model
#' y = response variable name in the fitted model
#' max_count = upper bound on the x-axis of the histograms
plot_predicted_fitted <- function(model, max_count = NULL){
  fitted.counts <- model$fitted.value
  response_name <- as.character(model$formula)[2]
  data1 <- data.frame(counts = c(fitted.counts, conf_counts[[response_name]]),
                      outcome = rep(conf_counts$outcome2, 2),
                      index_vte_type = rep(conf_counts$index_vte_type, 2),
                      los_total = rep(conf_counts$los_total, 2),
                      los_total_censored = rep(conf_counts$los_total_censored, 2),
                      n_comorb = rep(conf_counts$n_comorb, 2),
                      group = rep(c("Predicted", "Observed"), each=nrow(conf_counts)))
  
  if(!is.null(max_count)) data1 <- data1[data1$counts <= max_count, ]
  
  ggplot(data=data1, aes(x=counts, group=group, fill=group)) +
  geom_histogram(data=data1, alpha=0.3, position="identity", binwidth = 2) +
  facet_grid(~outcome)
}
```

## Poisson regression on the number of VTE hospitalizations by AC3Mo.
```{r fig.width=15,fig.height=4}
fit.n_hosp_vte <- zeroinfl(data=conf_counts, n_hosp_vte_prim_censored ~ outcome2 + offset(log(enroll_length)) | 1, dist = "poisson")
summary(fit.n_hosp_vte)
print("Exponentiated coefficients"); exp(coef((fit.n_hosp_vte)))

plot_predicted_fitted(model=fit.n_hosp_vte, max_count=25)
```

## Poisson regression on the number of non-VTE hospitalizations by AC3Mo.
```{r fig.width=15,fig.height=4}
fit.n_hosp_nonvte <- zeroinfl(data=conf_counts, n_hosp_prim_censored ~ outcome2 + offset(log(enroll_length)) | 1, dist = "poisson")
summary(fit.n_hosp_nonvte)
print("Exponentiated coefficients"); exp(coef((fit.n_hosp_nonvte)))

plot_predicted_fitted(model=fit.n_hosp_nonvte, max_count=20)
```


```{r fig.width=15,fig.height=4}
# negative binomial
fit.n_hosp_nonvte.negbin <- zeroinfl(data=conf_counts, n_hosp_nonvte_prim_censored ~ outcome2 + offset(log(enroll_length)) | 1, dist = "negbin")
summary(fit.n_hosp_nonvte.negbin)
print("Exponentiated coefficients"); exp(coef((fit.n_hosp_nonvte.negbin)))

plot_predicted_fitted(model=fit.n_hosp_nonvte.negbin, max_count=20)
```


```{r fig.width=15,fig.height=4}
# add index vte type
fit.n_hosp_nonvte <- zeroinfl(data=conf_counts, n_hosp_nonvte_prim_censored ~ outcome2 + offset(log(enroll_length)) | 1, dist = "poisson")
summary(fit.n_hosp_nonvte)
print("Exponentiated coefficients"); exp(coef((fit.n_hosp_nonvte)))

plot_predicted_fitted(model=fit.n_hosp_nonvte, max_count=20)
```



*Why is the null likelihood greater than the full likelihood?*
```{r}
fit.n_hosp_nonvte.null <- update(fit.n_hosp_nonvte, . ~ 1)
2 * (logLik(fit.n_hosp_nonvte) - logLik(fit.n_hosp_nonvte.null))
pchisq(2 * (logLik(fit.n_hosp_nonvte) - logLik(fit.n_hosp_nonvte.null)), df = 5, lower.tail = FALSE)
```


### Compare zero-inflated Poisson and usual Poisson models
```{r fig.width=15,fig.height=4}
summary(fit.n_hosp_nonvte.glm <- glm(data=conf_counts, n_hosp_nonvte_prim_censored ~ outcome2 + index_vte_type + n_comorb + los_total_censored + offset(log(enroll_length)), family = "poisson"))
# vuong(fit.n_hosp_nonvte.glm, fit.n_hosp_nonvte)

plot_predicted_fitted(model=fit.n_hosp_nonvte.glm, max_count=25)
```


```{r fig.width=15,fig.height=4}
fit1 <- zeroinfl(data=conf_counts, n_hosp_nonvte_prim_censored ~ outcome2 + index_vte_type + n_comorb + offset(log(enroll_length)) | 1, dist = "poisson")
summary(fit1)
plot_predicted_fitted(model=fit1, max_count=25)
```



# Obtain the number of distinct ICD-9 codes in medical data
```{r}
medical_info <- fread("../data/prog7_get_med.txt")

diag_names <- paste0("diag", 1:25)

icd9_3 <- medical_info[, (diag_names) := lapply(.SD, function(x) substr(x, 1, 3)), .SDcols = diag_names]

# unique across all columns?
```






