---
title: "checking for provider information"
output: slidy_presentation
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mengbing/Box Sync/OptumInsight_DataManagement/prep')

library(sas7bdat)
library(dplyr)
library(tidyr)
library(data.table)
library(knitr)
library(kableExtra)
```

## head(provider)

```{r}
provider <- data.table(fread("../data/prog12_prov.txt",
                       colClasses=c("character","character","Date","character",
                                    "character","numeric", "integer",
                                    "numeric","character","character", "character",
                                    "integer",
                                    "character","character", "integer", "character",
                                    "character","character","character", "character",
                                    "character","character")))
names(provider) <- tolower(names(provider))

knitr::kable(head(provider)) %>%
    kable_styling(latex_options = c("striped", "scale_down"), font_size = 14)
```


## 3 types

```{r echo=TRUE}
npi <- c("X1", "X2", "X3", "X3", "X3")
prov_unique <- c(0, NA, 34567, 24680, 0)
prov_cat <- c(0000, NA, 5018, 5957, 0000)
demo <- data.frame(npi, prov_unique, prov_cat)

knitr::kable(demo, format = "html", caption = "Three types of provider information structures") %>%
  kable_styling(full_width = F, latex_options = c("striped"), position = "center") %>%
  row_spec(1, background = "sandybrown") %>%
  row_spec(2, background = "palevioletred") %>%
  row_spec(3:5, background = "orangered") %>%
  column_spec(1:3, width = "4cm")
```




## NPI

### Histogram of anticoagulant fills with missing NPI
Missing NPI is defined as having values ""(empty), "L", or "LLLLLLLLLL".

```{r}
provider2 <- provider[, npi_miss := (npi2 %in% c("missing", "trivial"))]

# number of missing NPI
x_npi <- provider2[npi_miss==TRUE, .N]
x_npi

# total number of anticoagulant prescriptions
n_npi <- provider2[, .N]
n_npi

# proportion of patients having missing prov_unique
x_npi / n_npi
```


```{r}
prop_npi <- provider2[,sum(npi_miss)/.N, by=patid]
hist(prop_npi$V1, breaks = 15, xlab = "Proportion", main = "Proportion of missing NPI among anticoagulant prescriptions")
```


## Individual level provider information

### Proportion of anticoagulant fills with missing prov\_unique.

```{r}
provider2 <- provider2[, prov_miss := ( is.na(prov_unique))]
prop_pres <- provider2[,sum(prov_miss)/.N, by=patid]
hist(prop_pres$V1, breaks = 15, xlab = "Proportion", main = "Proportion of missing prov_unique among anticoagulant prescriptions")
```


## Histogram of at least one missing prov_unique.
```{r}
# number of patients having missing prov_unique
length(unique(prop_pres[V1!=0,patid]))

# total number of patients
length(unique(provider2$patid))

# proportion of patients having missing prov_unique
length(unique(prop_pres[V1!=0,patid])) / length(unique(provider2$patid))
```


```{r}
hist(prop_pres[V1!=0,V1], breaks = 15, xlab = "Proportion", main = "Proportion of at leat one missing prov_unique among anticoagulant prescriptions")
```



## One NPI corresponds to multiple prov_unique's
```{r}
p_unique <- provider2[, .N,by=.(patid, clmid, npi2)]

p_unique_multi <- p_unique[N>1,]

# number of NPI having more than 1 prov_unique
length(unique(p_unique_multi$npi2))

length(unique(p_unique_multi$patid))
```

```{r}
p_unique <- p_unique[, nobs := .N, , by=patid][, prop := N/nobs]
max(p_unique$prop)
hist(p_unique$prop, breaks = 5, main = "Proportion of NPI's having multiple prov_unique", xlab = "Proportion")
```



## Population level provider information

```{r}
# the number of AC prescriptions with npi having missing prov_unique
provider2[,sum(prov_miss)]

# total number of AC prescriptions of all patients
provider2[,.N]

# proportion of AC prescriptions with missing prov_unique
provider2[,sum(prov_miss)/.N]
```




## Example 1

Non-trivial NPI corresponds to a missing prov_unique

```{r echo = TRUE}
knitr::kable(provider[patid=="802666500415835"]) %>%
    kable_styling(latex_options = c("striped", "scale_down"), font_size = 14)
```



## Example 2

This patient got one drug in one day, but the corresponding NPI matches different prov_unique.

```{r echo = TRUE}
knitr::kable(provider[patid=="802666507718624"]) %>%
    kable_styling(latex_options = c("striped", "scale_down"), font_size = 14)
```


## Example 3

This patient did not have a valid prov_unique at either index AC or AC3Mo.

```{r echo = TRUE}
knitr::kable(provider[patid=="802666506295533"]) %>% 
    kable_styling(latex_options = c("striped", "scale_down"), font_size = 14)
```




## Example 4
This patient got one drug in one day with differnt NPI's, which match different prov_unique.

```{r echo = TRUE}
knitr::kable(provider[patid=="802666509497634"]) %>%
    kable_styling(latex_options = c("striped", "scale_down"), font_size = 14)
```







