---
title: "Multinomial logistic regressions on AC3Mo"
output:
  html_document: default
  html_notebook: deafult
---

## Summary:
At 3 months, people with "Other" and "Not captured" are removed. Remaining sample size = 12,652.

In sensitivity analysis, at 4 months people with "Other" and "Not captured" are removed. Remaining sample size = 10,134.

Note that patients used in the sensitivity analysis is a subset of the analysis at 3 months.


### About the covariates

The following argues why certain variables are not included in the regression model.

- In all SES variables including **race**, **education**, **federal poverty level**, **home ownership, occupation, income, networth**, and **division**, all NA values are coded as the same as the unknown level ("0" or "U").  
In the source data, race has values U, A, H, W, B, and "" (blank), where U means "Unknown". Networth range has values 0, 1, 2, 3, 4, 5, and "" (blank), where 0 means "Unknown". In the model, all blank values are filled in with the unknown level coding "U" or "0". In the end, race only has values  U, A, H, W, and B; networth range only has values 0,1,2,3,4,5. Variables having a level "Unknown" and having blank values are modified in this way. Such variables include race, education, federal poverty level, home ownership, occupation, income, networth range, and division.  
A patient may either have no NAs in any of these 8 SES variables, or have NAs in all of these SES variables. 375 patients (2.51%) have NAs in the SES variables.

- **Federal poverty level** only has 3 levels "Above 400% FPL", "Below 400% FPL", and "Unkown", and very few people (<=3) have "Below 400% FPL". Thus **fed\_poverty** is not used.

- **Home ownership** only has "probable homeowner" and "unknown". This does not provide useful information so is not used.

- **Occupation** has >75\% "unknown" and is thus not used.



Categories of some variables are combined, due to small cell counts which causes difficulty in fitting the model and yielding good parameter estimates.

Here is description of variables used in the model:

- **VTE_VTEName**: binary indicators of whether a patient has a particular VTE type (VTEName) in index\_vte\_type. It is derived from variable index\_vte\_type. Due to small counts of "IVC", "RV" and "PV", these three categories are combined into "ivc_rv_pv". Hence there are five such binary columns: "VTE_Lower.extremity.DVT", "VTE_Other", "VTE_Pulmonary.embolism", "VTE_Upper.extremity.DVT", and "VTE_ivc_rv_pv". For example, "VTE_Lower.extremity.DVT" = 1 if a patient has LE as one of index VTE types. This coding allows for multiple index VTE types.

- **vte\_history**: = 1 if a patient has history of VTE (ICD-9 code V12.51) within 12 months prior to index VTE date, and 0 if not.

- **Hospitalized**: = 1 if a patient was hospitalized within 28 days prior to index VTE date, and 0 if not.

- **malignancy_cancerName**: binary indicators of whether a patient has a particular cancer (cancerName) in malignancy. It is derived from cancer\_type. There are 8 malignancy types: Brain.CNS, Breast, Gastrointestinal, Genitourinary, Gynecologic, Hematologic, Lung, and Other. Example: malignancy_Brain.CNS = 1 if brain/CNS is one of the patient's index cancer types, and 0 if not. Malignancies follow the categorization in sheet "Table 1 Cleaned for Abstract" in "Optum Table 1 and 2 w regression 7-26-18.xlsx". This coding allows for multiple malignancies.

- **charlson_comorb_score**: Charlson comorbidity score.

- **Division**: categorical variable with 10 levels. "EAST NORTH CENTRAL" is the reference level.

- **Education2**: categorical variable. Due to small count in the category "Less than 12th Grade" ("A"), it is merged with level "High School Diploma" ("B") into in to a new level coded as "AB". **Education2** has four levels "AB", "C", "D", and "U", where "<Bachelor degree" ("C") is the reference level.

- **Income\_range**: categorical variable with 7 levels "0", ..., "6". "6" is the reference level.

- **ac3mo_copay_rng**: categorical variable with 7 levels [0, 10), [10, 30), [30, 50), [50, 100), [100, 250), [250, 500), and >= 500. The "Undefined" level is dropped because it only exists for Multiple/Unknown. "[0, 10)" is the reference level.

- **Product**: categorical variable with 6 levels. "HMO" is the referene level.

- **Age\_s**: continuous variable. Scaled age, defined as (age - mean of age) / standard deviation of age. 

- **Race**: categorical variable. It has levels "A", "B", "H", "W", and "U", where "W" (white) is the reference level.

- **Male**: = 1 if a patient is male, and 0 if female.





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/mengbing/Box Sync/OptumInsight_DataManagement/analysis/a1_proportions')

library(dplyr)
library(tidyr)  
library(data.table)
library(knitr)
library(kableExtra)
library(reshape2)
library(xlsx)
library(readxl)
library(mlogit)
library(nnet)
```


```{r echo=FALSE}
dat <- data.table(readRDS("../../data/analysis_data.rds"))

vars_not_wanted <- c("clmid", "fill_dt", "category", "brand_name", "gen_name", "copay", "copay_sum", "days_sup", "quantity", "strength", "npi")
dat1 <- unique(dat[,(vars_not_wanted) := NULL])

# UNKNOWN VALUE IN THESE VARIABLES ARE CODED AS "U"
ses1 <- c("education", "fed_poverty",  "occupation")
dat2 <- dat1[, (ses1) := lapply(.SD, function(x) ifelse(is.na(x), "U", x)), .SDcols=ses1]

# UNKNOWN VALUE IN THESE VARIABLES ARE CODED AS "0"
ses2 <- c("home_ownership", "networth_range", "income_range")
dat2[, (ses2) := lapply(.SD, function(x) ifelse(is.na(x), "0", x)), .SDcols=ses2]


# combine all doacs categories into doac
dat2$ac3mo2 <- with(dat2, ifelse(ac3mo %in% c("Apixaban", "Dabigatran", "Edoxaban", "Rivaroxaban"), "DOAC", ac3mo))
dat2[, ac3mo2 := as.factor(ac3mo2)]
dat2 <- within(dat2, ac3mo2 <- relevel(ac3mo2, ref = "LMWH"))

dat2$ac4mo2 <- with(dat2, ifelse(ac4mo %in% c("Apixaban", "Dabigatran", "Edoxaban", "Rivaroxaban"), "DOAC", ac4mo))
dat2[, ac4mo2 := as.factor(ac4mo2)]
dat2 <- within(dat2, ac4mo2 <- relevel(ac4mo2, ref = "LMWH"))


# unwrap combined index VTE types into columns of binary indicators
vte_info <- read_excel("../../data/prog8_vte.xlsx", sheet = "raw_VTE_POS")
vte_info2 <- unique(data.table(vte_info)[!is.na(VTE_type),.(patid, VTE_type)])
vte <- model.matrix(patid~VTE_type+0, data=vte_info2)
colnames(vte) <- gsub("type", "", colnames(vte))
vte2 <- data.frame(cbind(vte_info2[,1], vte))
vte3 <- aggregate(. ~ patid, vte2, sum)
dat2 <- merge(x=dat2, y=vte3, by="patid", all.x = TRUE)
rm(vte_info, vte_info2, vte, vte2, vte3)

# combine the 3 rarest VTE types
dat2[,VTE_ivc_rv_pv:= as.numeric((VTE_IVC | VTE_Portal.vein | VTE_Renal.vein))]

comorb_names <- colnames(dat2)[grep("comorbidity", colnames(dat2))]
dat2[, (comorb_names) := lapply(.SD, function(x) ifelse(is.na(x), 0, x)), .SDcols = comorb_names]
dat2[,comorb01 := !is.na(comorbidities)]
# count the number of comorbidities
dat2[, ("n_comorb") := rowSums(.SD), .SDcols=comorb_names]
# summary(dat2$n_comorb)

dat2[, race := ifelse(race=="", "U", race)]

antip_names <- colnames(dat2)[grep("antiplatelet", colnames(dat2))]
dat2[, (antip_names) := lapply(.SD, function(x) ifelse(is.na(x), 0, x)), .SDcols = antip_names]
dat2[, ("antiplatelet") := (rowSums(.SD) > 0), .SDcols=antip_names]

dat2[,c("race", "division", "product", "education", "fed_poverty","income_range", "networth_range", "occupation")] <- lapply(dat2[,c("race", "division", "product", "education", "fed_poverty","income_range", "networth_range", "occupation")], as.factor)

# pdf("a1_hist_n_comorbidities.pdf")
# hist(dat2$n_comorb, main="Histogram of number of comorbidities",
     # xlab="Number of comorbidities")
# dev.off()

# scaled age
dat2$age_s <- scale(dat2$age)
# center age by mean
dat2$age_c <- dat2$age - mean(dat2$age)

# combine the smallest education level to the next level
dat2[,education2 := as.factor(ifelse(as.character(education) %in% c("B", "A"), "AB", as.character(education)))]

# use the largest level as the reference level
dat2 <- within(dat2, education <- relevel(education, ref = "C"))
dat2 <- within(dat2, education2 <- relevel(education2, ref = "C"))
dat2 <- within(dat2, race <- relevel(race, ref = "W"))
dat2 <- within(dat2, income_range <- relevel(income_range, ref = "6"))
dat2 <- within(dat2, product <- relevel(product, ref = "HMO"))
dat2 <- within(dat2, ac3mo_copay_rng <- relevel(as.factor(ac3mo_copay_rng), ref="[0,10)"))
dat2 <- within(dat2, ac4mo_copay_rng <- relevel(as.factor(ac4mo_copay_rng), ref="[0,10)"))

# combine asian and hispanic for race
dat2[,race2 := as.factor(ifelse(as.character(race) %in% c("H", "A"), "AsianHispanic", as.character(race)))]

# combine regions for division
dat2 <- data.table(dat2 %>%
  mutate(division2 = case_when(
    division %in% c("WEST SOUTH CENTRAL", "WEST NORTH CENTRAL") ~ "WEST CENTRAL",
    division %in% c("EAST SOUTH CENTRAL", "EAST NORTH CENTRAL") ~ "EAST CENTRAL",
    division %in% c("SOUTH ATLANTIC", "MIDDLE ATLANTIC", "NEW ENGLAND") ~ "ATLANTIC_ENGLAND",
    division %in% c("MOUNTAIN", "PACIFIC") ~ "MOUNTAIN_PACIFIC",
    TRUE ~ "UNKNOWN"
  )))

# combine insurace type
dat2[,product2 := as.factor(ifelse(as.character(product) %in% c("EPO", "IND", "PPO"), "EPO_IND_PPO", as.character(product)))]
dat2 <- within(dat2, product2 <- relevel(product2, ref = "HMO"))

# obtain variable names for index VTE types
# all distinct VTE categories
vte_names0 <- colnames(dat2)[grep("VTE_", colnames(dat2))]
vte_names1 <- vte_names0[-which(vte_names0 %in% c("VTE_IVC","VTE_Portal.vein", "VTE_Renal.vein"))]


# Add combined malignancies
malignancy_names <- colnames(dat2)[grep("malignancy_", colnames(dat2))]


# remove people who have Other AC
data_ac3mo <- dat2[!ac3mo %in% c("Other", "Not captured"),]
data_ac3mo$ac3mo2 <- factor(data_ac3mo$ac3mo2)
data_ac3mo$ac3mo_copay_rng <- factor(data_ac3mo$ac3mo_copay_rng)

data_ac4mo <- dat2[!ac4mo %in% c("Other", "Not captured"),]
data_ac4mo$ac4mo2 <- factor(data_ac4mo$ac4mo2)
data_ac4mo$ac4mo_copay_rng <- factor(data_ac4mo$ac4mo_copay_rng)

round_pvalues <- function(x){
  if (x < 0.0001) return("<0.0001") else if (0.0001 <= x & x < 0.001)
    return("<0.001") else if (0.001 <= x) return(round(x, 3))
}

create_coefficients_table <- function(model, title_index){
  s <- summary(model)

  s_coef <- round(data.matrix(s$coefficients), 3)
  s_se <- round(data.matrix(s$standard.errors), 3)
  
  z <- s$coefficients/s$standard.errors
  # 2-tailed Wald z tests to test significance of coefficients
  p <- data.matrix(sapply((1-pnorm(abs(z), 0, 1))*2, round_pvalues))
  s_coef_exp <- round(exp(s_coef),3)
    
  coef_table <- t(matrix(paste0(s_coef, "±", s_se, " (", p, ")", "\\\n", s_coef_exp), nrow=3))
  colnames(coef_table) <- rownames(s_coef)
  rownames(coef_table) <- colnames(s_coef)
  title_name1 <- paste0("Model ", title_index, ". Results of multinomial logistic regression model. \\\n Output format: coefficient ± standard error (p-value from Wald test) \\\n \\\t \\\t \\\t \\\t exponentiated coefficient")
  kable(coef_table, format="html", caption = title_name1) %>%
    kable_styling(bootstrap_options = c("striped", "hover"))
  
}
```


## Multinomial logistic regression on AC at 3 months

### Model 1:
$\log{ \left( \frac{P(\text{AC3Mo = j})}{P(\text{AC3Mo = LMWH})} \right)} =$ vte\_VTENames + vte\_history + hospitalized + malignancy + charlson_comorb_score, where $j$ can be "Warfarin", "DOAC", "Unknown/Multiple".

i.e. LMWH is the reference level for AC3Mo.

```{r echo=FALSE}
fm1 <- as.formula(paste("ac3mo2~", paste(c(vte_names1), collapse="+"), "+ vte_history + hospitalized + ", paste(malignancy_names, collapse = "+"), "+ charlson_comorb_score"))
fit1 <- multinom(fm1, data=data_ac3mo, trace=FALSE)
create_coefficients_table(fit1, "1")
```



### Model 2:
$\log{ \left( \frac{P(\text{AC3Mo = j})}{P(\text{AC3Mo = LMWH})} \right)} =$ vte\_VTENames + vte\_history + hospitalized + malignancy + charlson_comorb_score + division, where $j$ can be "Warfarin", "DOAC", "Unknown/Multiple".

```{r echo=FALSE}
fm2 <- as.formula(paste("ac3mo2~", paste(c(vte_names1), collapse="+"), "+ vte_history + hospitalized + ", paste(malignancy_names, collapse = "+"), "+ charlson_comorb_score + division"))
fit2 <- multinom(fm2, data=data_ac3mo, trace=FALSE)
create_coefficients_table(fit2, "2")
```


## Model 3:
$\log{ \left( \frac{P(\text{AC3Mo = j})}{P(\text{AC3Mo = LMWH})} \right)} =$ vte\_VTENames + vte\_history + hospitalized + malignancy + charlson_comorb_score + division + education2 + income\_range + ac3mo\_copay\_range, where $j$ can be "Warfarin", "DOAC", "Unknown/Multiple".

```{r echo=FALSE}
fm3 <- as.formula(paste("ac3mo2~", paste(c(vte_names1), collapse="+"), "+ vte_history + hospitalized + ", paste(malignancy_names, collapse = "+"), "+ charlson_comorb_score + division + education2 + income_range + ac3mo_copay_rng"))
fit3 <- multinom(fm3, data=data_ac3mo, trace=FALSE)
create_coefficients_table(fit3, "3")
```



## Model 4:
$\log{ \left( \frac{P(\text{AC3Mo = j})}{P(\text{AC3Mo = LMWH})} \right)} =$ vte\_VTENames + vte\_history + hospitalized + malignancy + charlson_comorb_score + division + education2 + income\_range + ac3mo\_copay\_range + product, where $j$ can be "Warfarin", "DOAC", "Unknown/Multiple".

```{r echo=FALSE}
fm4 <- as.formula(paste("ac3mo2~", paste(c(vte_names1), collapse="+"), "+ vte_history + hospitalized + ", paste(malignancy_names, collapse = "+"), "+ charlson_comorb_score + division + education2 + income_range + ac3mo_copay_rng + product"))
fit4 <- multinom(fm4, data=data_ac3mo, trace=FALSE)
create_coefficients_table(fit4, "4")
```


## Model 5: An additional one to include variables not in the Excel
$\log{ \left( \frac{P(\text{AC3Mo = j})}{P(\text{AC3Mo = LMWH})} \right)} =$ vte\_VTENames + vte\_history + hospitalized + malignancy + charlson_comorb_score + division + education2 + income\_range + ac3mo\_copay\_range + product + age\_s + male + race, where $j$ can be "Warfarin", "DOAC", "Unknown/Multiple".

```{r echo=FALSE}
fm5 <- as.formula(paste("ac3mo2~", paste(c(vte_names1), collapse="+"), "+ vte_history + hospitalized + ", paste(malignancy_names, collapse = "+"), "+ charlson_comorb_score + division + education2 + income_range + ac3mo_copay_rng + product + age_s + male + race"))
fit5 <- multinom(fm5, data=data_ac3mo, trace=FALSE)
create_coefficients_table(fit5, "5")
```


**Notes on interpreting the results**

We focus on the largest model, Model 4.

- An example of interpreting the coefficients: The exponentiated coefficient of "VTE_Lower.extremity.DVT" for warfarin is 1.231. This means that the probability of AC3Mo being warfarin is 1.231 times the probability of AC3Mo being LMWH for a patient with LE as index VTE type, holding all other covariates constant.

- The model considers multiple index VTE types and multiple malignancies. For example, Let's consider a multiple VTE types case. The exponentiated coefficient of "VTE_Pulmonary.embolism" for warfarin is 0.834. Along with 1.231 from the previous point, the probability of AC3Mo being warfarin is 1.231\*0.834 = 1.027 times the probability of AC3Mo being LMWH for a patient with both LE and PE, holding all other covariates constant.

- Controlling for all other covariates, having any index VTE types decreases the relative probability of AC3Mo being DOAC compared to LMWH because all coefficients of DOAC are negative. In addition, having any of PE, UE, or ivc_rv_pv lowers the relative probability of AC3Mo being Warfarin compared to LMWH. For example,
for a patient with PE and UE simutaneously, the probability of AC3Mo being DOAC is 0.834\*0.915 = 0.763 times that being LMWH. Similarly, for a patient with the same condition, the probability of AC3Mo being warfarin is 0.88\*0.918 = 0.808 times that being LMWH.

- Patients with a VTE history are slightly less likely (0.786) to get DOAC at 3 months compared to LMWH, but are more likely (1.251) to get warfarin at 3 months than to get LMWH, holding other covariates constant. 

- Hospitalization decreases the probability of DOAC (0.845) or warfarin (0.75) at 3 months, compared to LMWH, holding other covariates constant.

- The coefficients of malignancy_cancerName are either positive or negative simultaneously for DOAC and Warfarin, except for malignancy_Hematologic. Hence having any malignancy type but hematologic changes the relative probability of DOAC and Warfarin, compared to LMWH, in the same direction. For a specific example, controlling for all other covariates, the probability of AC3Mo being DOACs is 0.404 times that of LMWH among patients having Brain.CNS as maglinancy, and the probability of AC3Mo being Warfarin is 0.498 times that of LMWH among patients having Brain.CNS as malignancy. 

- Per point increase in Charlson comorbidity score increases the relative probability of AC3Mo being DOAC compared to LMWH by 1.005 times, and that of Warfarin by 1.05 times. Both are very close to 1.

- The coefficients of division are either positive or negative simultaneously for DOAC and Warfarin, except for divisionSOUTH ATLANTIC and divisionUNKNOWN. Hence being in any region but SOUTH ATLANTIC and UNKNOWN changes the relative probability of DOAC and Warfarin, compared to LMWH, in the same direction. Specifically, people in EAST SOUTH CENTRAL are much more likely to have DOAC (2.487) and Warfarin (1.829) at 3 months compared to LMWH. On the other hand people in New England are much less likely to have DOAC (0.287) and Warfarin (0.476) at 3 months compared to LMWH, holding all other covariates constant. 

- Compared to patients having education level <Bachelor degree, having Bachelor degree + decreases the relative probability of AC3Mo being DOAC compared to LMWH by 0.773 times. Similarly, compared to patients having education level <Bachelor degree, having Bachelor degree + decreases the relative probability of AC3Mo being Warfarin compared to LMWH by 0.848 times, holding other covariates constant.

- Compared to patients in household income level \$100K+, having any other household income levels increases the relative probability of AC3Mo being Warfarin compared to LMWH because all coefficients for Warfarin are positive. On the other hand, having any other household income levels does not seem to change much the relative probability of AC3Mo being DOAC compared to LMWH because all coefficients are close to 0, no matter positive or negative.





# Multinomial logistic regression on AC at 4 months

### Model 1:
$\log{ \left( \frac{P(\text{AC4Mo = j})}{P(\text{AC4Mo = LMWH})} \right)} =$ vte\_VTENames + vte\_history + hospitalized + malignancy + charlson_comorb_score, where $j$ can be "Warfarin", "DOAC", "Unknown/Multiple".

i.e. LMWH is the reference level for AC4Mo.

```{r echo=FALSE}
fm1_4mo <- as.formula(paste("ac4mo2~", paste(c(vte_names1), collapse="+"), "+ vte_history + hospitalized + ", paste(malignancy_names, collapse = "+"), "+ charlson_comorb_score"))
fit1_4mo <- multinom(fm1_4mo, data=data_ac4mo, trace=FALSE)
create_coefficients_table(fit1_4mo, "1")
```



### Model 2:
$\log{ \left( \frac{P(\text{AC4Mo = j})}{P(\text{AC4Mo = LMWH})} \right)} =$ vte\_VTENames + vte\_history + hospitalized + malignancy + charlson_comorb_score + division, where $j$ can be "Warfarin", "DOAC", "Unknown/Multiple".

```{r echo=FALSE}
fm2_4mo <- as.formula(paste("ac4mo2~", paste(c(vte_names1), collapse="+"), "+ vte_history + hospitalized + ", paste(malignancy_names, collapse = "+"), "+ charlson_comorb_score + division"))
fit2_4mo <- multinom(fm2_4mo, data=data_ac4mo, trace=FALSE)
create_coefficients_table(fit2_4mo, "2")
```


## Model 3:
$\log{ \left( \frac{P(\text{AC4Mo = j})}{P(\text{AC4Mo = LMWH})} \right)} =$ vte\_VTENames + vte\_history + hospitalized + malignancy + charlson_comorb_score + division + education2 + income\_range + ac4mo\_copay\_range, where $j$ can be "Warfarin", "DOAC", "Unknown/Multiple".

```{r echo=FALSE}
fm3_4mo <- as.formula(paste("ac4mo2~", paste(c(vte_names1), collapse="+"), "+ vte_history + hospitalized + ", paste(malignancy_names, collapse = "+"), "+ charlson_comorb_score + division + education2 + income_range + ac4mo_copay_rng"))
fit3_4mo <- multinom(fm3_4mo, data=data_ac4mo, trace=FALSE)
create_coefficients_table(fit3_4mo, "3")
```



## Model 4:
$\log{ \left( \frac{P(\text{AC4Mo = j})}{P(\text{AC4Mo = LMWH})} \right)} =$ vte\_VTENames + vte\_history + hospitalized + malignancy + charlson_comorb_score + division + education2 + income\_range + ac4mo\_copay\_range + product, where $j$ can be "Warfarin", "DOAC", "Unknown/Multiple".

```{r echo=FALSE}
fm4_4mo <- as.formula(paste("ac4mo2~", paste(c(vte_names1), collapse="+"), "+ vte_history + hospitalized + ", paste(malignancy_names, collapse = "+"), "+ charlson_comorb_score + division + education2 + income_range + ac4mo_copay_rng + product"))
fit4_4mo <- multinom(fm4_4mo, data=data_ac4mo, trace=FALSE)
create_coefficients_table(fit4_4mo, "4")
```





```{r echo=FALSE, eval=FALSE}
# running ANOVA to compare mean ages across groups
fit_age <- aov(age ~ ac3mo2, data=data_ac3mo)
summary(fit_age)

# running ANOVA to compare sex proportions across groups
fit_sex <- aov(male ~ ac3mo2, data=data_ac3mo)
summary(fit_sex)
```











